Sound Automated Synthesis of Digital Controllers

### 2016-10-08 paper/paper.tex
New figures

In this paper, however, we explore the automatic synthesis of 
{\em correct-by-construction} dgital controllers, which 
reduces the deveopment cost and design time for a number of models.  
In particular, we are interested in synthesising controllers 
for {\em hybrid closed-loop systems}, {\it i.e.}, embedded controllers along with a model of their
physical environment (the plant), which are {\em stable}.  Due to the complexity
of such systems, we focus on linear models with known configurations
to perform parametric synthesis of the controller dynamics.

@@

%\begin{figure}
%\centering
%\resizebox{.35\textwidth}{!}{
% \begin{tikzpicture}[scale=0.3,-,>=stealth',shorten >=.2pt,auto,
%     semithick, initial text=, ampersand replacement=\&,]

%  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.5cm, column sep=.5cm] {
%   \coordinate (aux0);
%   \& \node[circle] (circle) {}; 
%   \& \node[fill=yellow!20] (da) {{\sc D/A}}; 
%   \& \node[fill=yellow!20] (p) {{\sc P}};
%   \& \node[fill=yellow!20] (ad) {{\sc A/D}}; 
%   \& \coordinate (aux1);\\
%   \& \coordinate (aux3); 
%   \&
%   \&
%   \node[fill=yellow!20] (c) {C};
%   \& 
%   \& \coordinate (aux2);\\
%  };

%  \path[->] (aux0) edge (circle.west);
%  \path  
%   (circle.east) edge (da.west)
%   (da.east) edge (p.west)
%   (p.east) edge (ad.west)
%   (ad.east) edge (aux1.west)
%   (aux1.south) edge (aux2.north)
%   (aux2.west) edge (c.east)
%   (c.west) edge (aux3.east); 
%  \path[->]  (aux3.north) edge (circle.south);
% \end{tikzpicture}
%}
% \caption{Closed-Loop System. \label{fig:closed-system}}
%\end{figure}

@@

This intermediate ANSI-C code model $(\tilde C(z),\mathfrak{G}(z),\phi)$
containing a specification $\phi$ for the verification properties ({\it i.e.}, robust stability)
is translated into SAT/SMT formulae by a back-end model-checking tool ({\it e.g.},
CBMC~\cite{ClarkeKL04} or ESBMC~\cite{CordeiroFM12}).
If any property violation is found, then DSVerifier reports a
counterexample, which contains system inputs or parametric deviations
$\Delta{G}$ that lead the system to a failure.  A successful verification
result is reported if the system is safe with respect to $\phi$.  In
particular, the stability verification is complete and sound, since it does
not depend on system inputs and outputs.
In the case of stability, the specification $\phi$ consists in a number of assumptions
on the polynomial coefficients, following Jury's Criteria, as well as the restrictions
set in the representation of these coefficients by the function $f$ (cf. Eq. \eqref{FWL_tf}).

@@

@@

\paragraph{General architecture of the program synthesiser}
%
%% \red{[I would recommend to clarify the domain of definition of $\sigma$, particularly over the set where $x$ ranges over. 
%% (Be cognisant that control audience might be unfamiliar with the cegis architecture.) ]}
% 
The input specification provided to the program synthesiser is of the form
$\exists \vec{P} . \forall \vec{x} \in \text{\sc inputs} . \sigma(\vec{x}, \vec{P})$
where $\vec{P}$ ranges over functions, $\vec{x}$ ranges over
ground terms and $\sigma$ is a quantifier-free formula.
We interpret the gound terms over some finite domain $\mathcal{D}$.

The design of our synthesiser is given in
Fig.~\ref{fig:CEGIS} and consists of two phases, {\sc
Synthesise} and {\sc Verify}, which interact via a finite set of test vectors {\sc inputs} 
that is updated incrementally.
Given the aforementioned specification $\sigma$, the {\sc synth} procedure 
tries to find an existential witness $P$
that satisfies the specification $\sigma(x, P)$ for all $x$.
%
If {\sc synthesise} succeeds in finding a witness $P$, this witness is a
candidate solution to the full synthesis formula.  We pass this candidate
solution to {\sc verify}, which determines whether it does satisfy
the specification on all inputs by checking satisfiability of the following 
verification formula:
%
$\exists x . \lnot \sigma(x, P)$

If this formula is unsatisfiable, the candidate solution is in fact a
solution to the synthesis formula and so the algorithm terminates. 
Otherwise, the witness $x$ is an input on which the candidate solution fails
to meet the specification.  This witness $x$ is added to the {\sc inputs}
set, and the loop iterates again.  It is worth noting that each iteration of the
loop adds a new input to the finite set $\text{\sc inputs}$  being used for synthesis.  
Given that the full set of inputs $\mathcal{D}$ 
%\blue{[which by now should have a name, as suggested to be introduced above]} 
is finite, this means that the refinement loop
can only iterate a finite number of times.

@@

We use this method to check the stability in the $z$-domain for the characteristic polynomial $S(z)$ of 
the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$, where $Y(z)$ is the 
closed-loop system output and U(z) is the controller output. 
%% \red{[unclear: what matrix? We need to clarify the relationship between the char. poly. $S(z)$ and the later quantities $\Delta N_{G}(z), \Delta D_{G}(z)$ and corresponding controller's structure.]}.
%
We consider the following form for $S(z)$:
%
$$
S(z) = a_0Z^N+a_1z^{N-1}+\cdots+a_{N-1}z+a_N=0, a_0\neq0
$$
%
%\cdsay{is this for the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$?}

Assuming that $\hat{G}(z) = \frac{\hat{N_G}(z)}{\hat{D_G}(z)}$
and $\hat{C}(z) = \frac{\hat{N_C}(z)}{\hat{D_C}(z)}$, then:
$$S(z) = FWL[N_C(z)] \times \hat{N_G}(z) + FWL[D_C(z)] \times \hat{D_G}(z)$$

@@

In the {\sc verify} phase, if the system is unstable for the current candidate solution,
{\it i.e.}, if $\neg \phi_{stability}$ is satisfiable for $S(z)$, 
then we obtain a counterexample denoted by 
values for each coefficients of $\Delta N_{G}(z)$ and $\Delta D_{G}(z)$, 
%\red{[see my comment above - these quantities pop up too abruptly.]},
which make the closed-loop system unstable. 
Note that, similar to the rest of the paper, we use the notation 
$\hat{N_G}(z) = N_G(z) + \Delta N_G(z)$ and 
$\hat{N_C}(z) = N_C(z) + \Delta N_C(z)$.
This uncertainty is added to 
the set {\sc inputs} such that, in the subsequent {\sc synthesise} phase, 
we obtain a candidate solution consisting 
of a controller $D_C$ and $N_C$, which makes the closed-loop 
system stable for all the uncertainties accumulated in {\sc inputs}.

\paragraph{Program synthesis for the running example}
We will illustrate our synthesis approach on the running example 
in Sec.~\ref{sec:running-ex}.
We start with a candidate solution that has all coefficients 0:
%\red{[Be more explicit, write out spurious candidate solution with zero coefficients. KISS - you don't want to lose the control audience here.]}.

$$
\begin{array}{ll}
N_C(z) {=} 0z^2{+}0z{+}0\\
D_C(z) {=} 0z^2{+}0z{+}0\\
\end{array}
$$

In the first {\sc verify} phase we obtain the following counterexample:
$$
\begin{array}{ll}
\Delta N_G(z) = 0.02634525299072265625\\
\Delta D_G(z) = 0.9979267120361328125z-0.99768829345703125
\end{array}
$$

@@

In the second {\sc verify} phase, we check whether the current candidate
solution above is the final solution. It is not and we get a new counterexample:
$$
\begin{array}{ll}
\Delta N_G(z) = 0.02644824981689453125\\
\Delta D_G(z) {=} 1.00053310394287109375z{-}1.0014629364013671875
\end{array}
$$
%\blue{[Nice, now I'm in the groove. Next unfolds well.] }
After adding this new counterexample to the set of {\sc inputs}, in the 
next {\sc synthesise} phase, we get a new candidate solution that holds
for both uncertainties in {\sc Inputs}:

>>>>

In this paper, however, we explore the automatic synthesis of 
{\em correct-by-construction} dgital controllers, which 
reduces the deveopment cost and design time for a number of models.  
In particular, we are interested in synthesising controllers 
for {\em hybrid closed-loop systems}, {\it i.e.}, embedded controllers along with a model of their
physical environment, the plant, as shown in
Fig.~\ref{fig:closed-system}, that are {\em stable}.  Due to the complexity
of such systems, we focus on linear models with known configurations
to perform parametric synthesis of the controller dynamics.

@@

\begin{figure}
\centering
\resizebox{.35\textwidth}{!}{
 \begin{tikzpicture}[scale=0.3,-,>=stealth',shorten >=.2pt,auto,
     semithick, initial text=, ampersand replacement=\&,]

  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.5cm, column sep=.5cm] {
   \coordinate (aux0);
   \& \node[circle] (circle) {}; 
   \& \node[fill=yellow!20] (da) {{\sc D/A}}; 
   \& \node[fill=yellow!20] (p) {{\sc P}};
   \& \node[fill=yellow!20] (ad) {{\sc A/D}}; 
   \& \coordinate (aux1);\\
   \& \coordinate (aux3); 
   \&
   \&
   \node[fill=yellow!20] (c) {C};
   \& 
   \& \coordinate (aux2);\\
  };

  \path[->] (aux0) edge (circle.west);
  \path  
   (circle.east) edge (da.west)
   (da.east) edge (p.west)
   (p.east) edge (ad.west)
   (ad.east) edge (aux1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (c.east)
   (c.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
 \caption{Closed-Loop System. \label{fig:closed-system}}
\end{figure}

@@

This intermediate ANSI-C code model $(\tilde C(z),\mathfrak{G}(z),\phi)$
containing a specification $\phi$ for the verification properties ({\it i.e.}, robust stability)
is translated into SAT/SMT formulae by a back-end model-checking tool ({\it e.g.},
CBMC~\cite{ClarkeKL04} or ESBMC~\cite{CordeiroFM12}).
If any property violation is found, then DSVerifier reports a
counterexample, which contains system inputs or parametric deviations
$\Delta{G}$ that lead the system to a failure.  A successful verification
result is reported if the system is safe with respect to $\phi$.  In
particular, the stability verification is complete and sound, since it does
not depend on system inputs and outputs.
In the case of stability, the specification $\phi$ consists in a number of assumptions
on the polynomial coefficients, following Jury's Criteria, as well as the restrictions
set in the representation of these coefficients by the function $f$ (Eq. \ref{FWL_tf}).

@@

\begin{figure}[ht]
\centering

  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.6cm, column sep=.3cm] {
    \node[draw=none] (r) {$R(z)$};
%   \& \coordinate (aux0);
   \& \node[circle,add={-}{+}{}{}] (circle) {};
   \node[draw=none] (ez) at ([xshift=1cm,yshift=.15cm]circle)  {$e(z)$};
   \& \node[rectangle,draw,
	minimum width=1cm,
	minimum height=1cm,
        label=\textbf{Controller}] (cz) {{\sc C(z)}};
   \node[draw=none] (ud) at ([xshift=1cm,yshift=.15cm]cz)  {$U(z)$};
   \& \node[rectangle,draw,
	minimum width=1cm,
	minimum height=1cm,
        label=\textbf{Plant}] (gs) {{\sc G(z)}};
   \node[draw=none] (y) at ([xshift=1cm,yshift=.15cm]gs)  {$Y(z)$};

   \& \node[circle,add={}{+}{+}{}] (circle2) {};
%   \node[draw=none] (tyz) at ([xshift=-1cm,yshift=.15cm]circle2)  {$Y(z)$};
   \node[draw=none] (nu) at ([yshift=1cm]circle2)  {$\nu(z)$};
   \& \coordinate (aux1);
   \& \node[draw=none] (yz) {$\tilde Y(z)$};\\
   \& \coordinate (aux3); 
   \&
   \& 
   \& 
   \& \coordinate (aux2);\\
  };

%  \path[->] (v1) edge (q1.north);
%  \path[->] (v2) edge (q2.north);
  \path[->] (r) edge (circle.west);
  \path[->] (nu) edge (circle2.north);
  \path[->] (aux1) edge (yz);
  \path  
   (circle.east) edge (cz)
   (cz.east) edge (gs.west)
   (gs.east) edge (circle2.west)
   (circle2.east) edge (aux1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
 \caption{Sampled System. \label{fig:sampledsystem}}
\end{figure}

%
%

\begin{figure}[ht]
\centering

  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.6cm, column sep=.3cm] {
    \node[draw=none] (r) {$R(z)$};
%   \& \coordinate (aux0);
   \& \node[circle,add={-}{+}{}{}] (circle) {};
   \node[draw=none] (ez) at ([xshift=1cm,yshift=.15cm]circle)  {$e(z)$};
   \& \coordinate (aux4);
   \& \node[rectangle,draw,
	minimum width=1cm,
	minimum height=1cm,
        label=\textbf{Controller}] (cz) {{\sc C(z)}};
   \node[draw=none] (ud) at ([xshift=1cm,yshift=.15cm]cz)  {$U(z)$};
   \& \node[rectangle,draw,
	minimum width=1cm,
	minimum height=1cm,
        label=\textbf{Plant}] (gs) {{\sc G(z)}};
   \node[draw=none] (y) at ([xshift=1cm,yshift=.15cm]gs)  {$Y(z)$};
   \& \node[circle,add={+}{+}{}{}] (circle2) {};
%   \node[draw=none] (tyz) at ([xshift=-1cm,yshift=.15cm]circle2)  {$Y(z)$};
%   \node[draw=none] (nu) at ([yshift=1cm]circle2)  {$\nu(z)$};
   \& \coordinate (aux1);
   \& \node[draw=none] (yz) {$\tilde Y(z)$};\\
   \& 
   \& \coordinate (aux5); 
   \& \node[rectangle,draw,
	minimum width=1cm,
	minimum height=1cm] (deltag) {{\sc $\Delta G(z)$}};
   \&
   \& \coordinate (aux6);\\
   \& \coordinate (aux3); 
   \&
   \& 
   \& 
   \&
   \& \coordinate (aux2);\\
  };

%  \path[->] (v1) edge (q1.north);
%  \path[->] (v2) edge (q2.north);
  \path[->] (r) edge (circle.west);
%  \path[->] (nu) edge (circle2.north);
  \path[->] (aux1) edge (yz);
  \path  
   (aux4.north) edge (aux5.south)
   (aux5.east) edge (deltag.west)
   (deltag.east) edge (aux6.west)
   (aux6.south) edge (circle2.south)
   (circle.east) edge (cz)
   (cz.east) edge (gs.west)
   (gs.east) edge (circle2.west)
   (circle2.east) edge (aux1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
 \caption{Sampled System. \label{fig:sampledsystem}}
\end{figure}

@@

\paragraph{General architecture of the program synthesiser}
%
\red{[I would recommend to clarify the domain of definition of $\sigma$, particularly over the set where $x$ ranges over. 
(Be cognisant that control audience might be unfamiliar with the cegis architecture.) ]}
% 
The design of our synthesiser is given in
Fig.~\ref{fig:CEGIS} and consists of two phases, {\sc
Synthesise} and {\sc Verify}, which interact via a finite set of test vectors {\sc inputs} \blue{that is updated incrementally}.
The {\sc synth} procedure tries to find an existential witness $P$
that satisfies the specification $\sigma(x, P)$ for all $x$, namely:
%
$\exists P . \forall x \in \text{\sc inputs} . \sigma(x, P)$

If {\sc synthesise} succeeds in finding a witness $P$, this witness is a
candidate solution to the full synthesis formula.  We pass this candidate
solution to {\sc verify}, which determines whether it does satisfy
the specification on all inputs by checking satisfiability of the following 
verification formula:
%
$\exists x . \lnot \sigma(x, P)$

If this formula is unsatisfiable, the candidate solution is in fact a
solution to the synthesis formula and so the algorithm terminates. 
Otherwise, the witness $x$ is an input on which the candidate solution fails
to meet the specification.  This witness $x$ is added to the {\sc inputs}
set, and the loop iterates again.  It is worth noting that each iteration of the
loop adds a new input to the \red{finite} set $\text{\sc inputs}$  being used for synthesis.  If
the full set of inputs \blue{[which by now should have a name, as suggested to be introduced above]} is finite, this means that the refinement loop
can only iterate a finite number of times.

@@

We use this method to check the stability in the $z$-domain for the characteristic polynomial $S(z)$ of 
the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$ \red{[unclear: what matrix? We need to clarify the relationship between the char. poly. $S(z)$ and the later quantities $\Delta N_{G}(z), \Delta D_{G}(z)$ and corresponding controller's structure.]}. 
We consider the following form for $S(z)$:
%
$$
S(z) = a_0Z^N+a_1z^{N-1}+\cdots+a_{N-1}z+a_N=0, a_0\neq0
$$
%
%\cdsay{is this for the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$?}

@@

In the {\sc verify} phase, if the system is unstable for the current candidate solution,
{\it i.e.}, if $\neg \phi_{stability}$ is satisfiable for $S(z)$, 
then we obtain a counterexample denoted by 
values for each coefficients of $\Delta N_{G}(z)$ and $\Delta D_{G}(z)$ \red{[see my comment above - these quantities pop up too abruptly.]},
which make the closed-loop system unstable. This uncertainty is added to 
the set {\sc inputs} such that, in the subsequent {\sc synthesise} phase, 
we obtain a candidate solution consisting 
of a controller $D_C$ and $N_C$, which makes the closed-loop 
system stable for all the uncertainties accumulated in {\sc inputs}.

\paragraph{Program synthesis for the running example}
We will illustrate our synthesis approach on the running example 
in Sec.~\ref{sec:running-ex}.
We start with a candidate solution that has all coefficients 0 \red{[Be more explicit, write out spurious candidate solution with zero coefficients. KISS - you don't want to lose the control audience here.]}.
In the first {\sc verify} phase we obtain the following counterexample:
$$
\begin{array}{ll}
\Delta N_G(z) = 0.02634525299072265625\\
\Delta D_G(z) = 0.9979267120361328125z-0.99768829345703125
\end{array}
$$

@@

In the second {\sc verify} phase, we check whether the current candidate
solution above is the final solution. It is not and we get a new counterexample:
$$
\begin{array}{ll}
\Delta N_G(z) = 0.02644824981689453125\\
\Delta D_G(z) {=} 1.00053310394287109375z{-}1.0014629364013671875
\end{array}
$$
\blue{[Nice, now I'm in the groove. Next unfolds well.] }
After adding this new counterexample to the set of {\sc inputs}, in the 
next {\sc synthesise} phase, we get a new candidate solution that holds
for both uncertainties in {\sc Inputs}:
